---
title: "Install"
format:
  html:
    toc: true
---

<style>
  body {
    background-repeat: repeat-y;
    background-size: cover;
    background-attachment: fixed;
    background-position: top center;
    position: relative;
    z-index: 0;
  }

  .background-overlay {
    position: fixed;
    top: 0;
    left: 0;
    width: 100%;
    height: 100%;
    background-color: rgba(255, 255, 255, 0.85);
    z-index: -1;
  }
</style>

## Install the package

You can install the latest development version of `fungaR` package from [GitHub](https://github.com/DBOSlab/fungaR) as follows:

```{r}
#| eval: false

if (!requireNamespace("BiocManager", quietly = TRUE))
install.packages("BiocManager")

# Install the development version of fungaR from GitHub,
# together with its required dependencies
BiocManager::install("DBOSlab/fungaR", dependencies = TRUE)
```

:::{.callout-caution}
`fungaR` requires R (>= 4.1.0). If you already have a recent R installation, you can skip straight to the install step above.
:::

```{r}
#| eval: false

library(fungaR)
```

## Quick Start

Once installed, a good first step is to check which Flora e Funga do Brasil dataset versions are currently published:

```{r}
#| eval: false

funga_version()
```

Most functions — including `funga_records()`, `funga_search()`, and `funga_get_children_taxa()` — download and parse the FFB dataset automatically the first time you call them, and reuse the same local cache on subsequent calls, so you rarely need to call `funga_download()`/`funga_parse()` yourself unless you want to inspect the raw data directly. Note that FFB publishes plants and fungi together in one combined release; `funga_parse()` is what subsets everything down to the Fungi kingdom.

From there, browse the [How-To Articles](/articles/index.qmd) to learn how to download and parse checklist data, filter records by taxonomic or geographic criteria, resolve your own species names against the checklist, explore the taxonomic hierarchy, and find species and state records missing from FFB.
